Distortion-corrected kernel density estimate on Riemannian manifolds

### 2022-07-15 paper/paper.Rmd
More on DC_KDE

However, in application, we usually find the observations lying on a $d$-dimensional manifold embedded in a much higher $s$-dimensional space, where $d \ll s$. Instead of sampling from the undering manifold $M$, we apply manifold learning algorithms to reduce dimension and find a $d$-dimensional representation of the manifold $M$.
Notice that this representation is not necessarily isometry to $M$, and the underlying manifold is usually unknown in real data sets. Then the density estimator in Equation \@ref(eq:denriem) proposed by @Pelletier2005-vu no longer applies because of the unknown geodesic distance $d_g$ and the volume density function $\theta_{\pmb{X}_i}(\pmb{p})$.
Motivated on the estimator in \@ref(eq:denriem), we propose a distortion-corrected kernel density estimator (DC-KDE) of $f$, same as \@ref(eq:denestimator0), with the manifold embedding $\pmb{y}_i, i = 1, 2, \dots, N$ to be 
\begin{equation}
\label{eq:denestimator}
\hat{f}(\pmb{p}) = \frac{1}{N} \sum_{i=1}^{N} \frac{1}{r^d} \frac{|\pmb{H}(\pmb{p})|}{|\pmb{H}(\pmb{y}_i)|} K\bigg( \frac{(\pmb{p} - \pmb{y}_i)^\prime H^{-1}(\pmb{y}_i) (\pmb{p} - \pmb{y}_i)}{r} \bigg),
\end{equation} 
where $\pmb{H}(\pmb{y}_i), i = 1, 2, \dots, N$ is the pointwise variable Riemannian matrix and it is positive definite with dimension $d\times d$, $\frac{|\pmb{H}(\pmb{p})|}{|\pmb{H}(\pmb{y}_i)|}$ is the estimated volume density function when using the coordinates from any manifold learning mapping, $(\pmb{p} - \pmb{y}_i)^\prime \pmb{H}^{-1}(\pmb{y}_i) (\pmb{p} - \pmb{y}_i)$ is the estimated geodesic distance with the distortion corrected by the Riemannian metric at embedding point $\pmb{y}_i \in \mathbb{R}^d$. Then the only problem left is the estimation of the Riemannian metric at each embedding point.

>>>>

However, in application, we usually find the observations lying on a $d$-dimensional manifold embedded in a much higher $s$-dimensional space, where $d \ll s$. Instead of sampling from the underlying manifold $M$, we could only sample from the high-dimensional data and apply dimension reduction techniques such as manifold learning to find a $d$-dimensional representation of $M$.
Notice that this representation is not necessarily isometry to $M$, and the underlying manifold is usually unknown in real data sets. We could use the low-dimensional embedding as an approximation of the true manifold $M$. However, the pointwise distortion induced by the manifold learning process needs to be taken care of to accurately estimate the density of the manifold. Therefore, the density estimator in Equation \@ref(eq:denriem) no longer applies because of the unknown geodesic distance $d_g$ and the volume density function $\theta_{\pmb{X}_i}(\pmb{p})$.
Further, the bandwidth matrix should be adaptive to each sample point to take into account the distortion.
Motivated by the estimator in \@ref(eq:denriem), we restate the proposed distortion-corrected kernel density estimator (DC-KDE) of $f$, same as \@ref(eq:denestimator0), based on the manifold embedding $\pmb{y}_i, i = 1, 2, \dots, N$,
\begin{equation}
\label{eq:denestimator}
\hat{f}(\pmb{p}) = \frac{1}{N} \sum_{i=1}^{N} \frac{1}{r^d} \frac{|\pmb{H}(\pmb{p})|}{|\pmb{H}(\pmb{y}_i)|} K\bigg( \frac{(\pmb{p} - \pmb{y}_i)^\prime H^{-1}(\pmb{y}_i) (\pmb{p} - \pmb{y}_i)}{r} \bigg),
\end{equation}
where $\pmb{H}(\pmb{y}_i), i = 1, 2, \dots, N$ is the pointwise variable Riemannian matrix and it is positive definite with dimension $d\times d$, $\frac{|\pmb{H}(\pmb{p})|}{|\pmb{H}(\pmb{y}_i)|}$ is the estimated volume density function when using the coordinates from any manifold learning mapping, $(\pmb{p} - \pmb{y}_i)^\prime \pmb{H}^{-1}(\pmb{y}_i) (\pmb{p} - \pmb{y}_i)$ is the estimated geodesic distance with the distortion corrected by the Riemannian metric at embedding point $\pmb{y}_i \in \mathbb{R}^d$. Then the only problem left is the estimation of the Riemannian metric at each embedding point.
